---
title: "Data analysis in R"
output: html_notebook
---
Assign a value to a variable
```{r}
x<-1
x
```
It can also be printed in console

```{r}
print(x)
```
To assign a character value
```{r}
msg<-"hello"
msg
```
To assign a vector
```{r}
x1<-1:20
x1
```
To use integers
```{r}
int<-2L
int
```
To use infinity
```{r}
int/Inf
```
To concatenate values
```{r}
c(0.5, 0.6)
```
To change values to logic
```{r}
as.logical(x1)
```
To create a list (It can contain values of different types)
```{r}
list_x<-list(1,"a", TRUE, 1+4i, 2L)
list_x
```
To create an empty matrix
```{r}
mat<-matrix(nrow=2, ncol=3)
mat
```
To see the dimension of the matrix created
```{r}
dim(mat)
```
To see the attributes of the matrix. In this case its dimension
```{r}
attributes(mat)
```
To fill a matrix
```{r}
mat<-matrix(1:6, nrow=2, ncol=3)
mat
```
To change the format of the matrix changing the dimension as attribute
```{r}
dim(mat)<-c(3,2)
mat
```
To concatenate vectors by rows 
```{r}
x2<-1:3
y2<-7:9
rbind(x2,y2)
```
To concatenate vectors by columns
```{r}
cbind(x2,y2)
```
To use factors to label the results in the table
```{r}
fac<-factor(c("yes", "yes", "no", "yes"))
fac
```
To see the factor as table
```{r}
table(fac)
```
To find null values
```{r}
is.na(fac)
```
To find nan values
```{r}
is.nan(fac)
```
To create a dataframe in R
```{r}
df<-data.frame(id=1:4, target=c(T,T,F,F))
df
```
To see the number of rows
```{r}
nrow(df)
```
To see the number of columns
```{r}
ncol(df)
```
To see the names of the columns
```{r}
names(df)
```
To create a list with names
```{r}
list_x1<-list(a=1,b=2,c=3)
list_x1
```
To create a list with names
```{r}
mat2<-matrix(1:4,nrow=2, ncol=2)
dimnames(mat2)<-list(c("a","b"),c("c","d"))
mat2
```
